Response to Ian's comments on $\metaprot$ to $\minifed$

### 2023-11-05 popl24/scratch.tex
checkpoint

{\footnotesize
\begin{verbatimtab}
  owl(gid) { { k = flip[2,"gate:" || gid || ".k"]; p = flip[2,"gate:" || gid || ".p"] } }

  encode(sa, sb) {
    let owl1 = owl(sa) in
    let owl2 = owl(sb) in
    v[1,gate: || sa || 1.k] := OT(s[1,sa],owl1.k,(not owl1.k));
    v[1,gate: || sa || 1.p] := OT(s[1,sa],owl1.p,(not owl1.p));
    v[1,gate: || sb || 2.k] := select(s[2,sb],owl2.k,(not owl2.k));
    v[1,gate: || sb || 2.p] := select(s[2,sb],owl2.p,(not owl2.p));
    { wv1 = { k = v[1,"gate:" || sa || "1.k"]; p = v[1,"gate:" || sa || "1.p"] };
      wv2 = { k = v[1,"gate:" || sb || "2.k"]; p = v[1,"gate:" || sb || "2.p"] } }
  }

  andgg(g, ga, gb, wva, wvb) { andgate(g, ga, gb); evalgate(g, wva, wvb) }

  decode(g, wv) { garbledecode(wv.g); evaldecode(v) }
\end{verbatimtab}
}

{\footnotesize
\begin{verbatimtab}
  let secrets = encode("s1","s2") in v[0,output] := decode(andgg("0", secrets.wv1, secrets.wv2)) \end{verbatimtab}
}

{\footnotesize
\begin{verbatimtab}
  owl    :   string(gid)

  encode :   string(sa) * string(sb)

  andgg  :  string(g) * { k : bool[1]; p : bool[1] } *  { k : bool[1]; p : bool[1] }

  decode :  string(g) * { k : bool[1]; p : bool[1] }
\end{verbatimtab}
}

>>>>

\documentclass[acmsmall,screen,review]{acmart}

\usepackage{amsmath}
\usepackage{amstext}
\usepackage{fp-frame}
\usepackage[latin1]{inputenc}
\usepackage{mathpartir}
\usepackage{fancyvrb}
\usepackage{moreverb}
\usepackage{stmaryrd}
\usepackage{enumerate}
\usepackage{thmtools,thm-restate}
\usepackage{comment}

\input{macros}

\begin{document}

\title{Response to Ian's comments on $\metaprot$ to $\minifed$}

\maketitle

\section{Model (Over)simplifications}

Ian comments that reviewers may push back on (1) our definition of OT
(actually borrowed from \cite{barthe2019probabilistic}) relies on a
trusted 3rd party for shared randomness, and (2) reviewers may push
back on our linear program and reasoning models that don't capture
concurrency that is a common element of MPC settings.

In fact, these were both strong criticisms of the POPL reviewers.
They are definitely valid criticisms, but I think we'd need to enrich
our framework to add them. For example, we could add a notion of
public key encryption as a language primitive and encode semi-honest
OT using the protocol described in \cite{evans2018pragmatic}, unless
there's a way to express public key cryptography in the binary field
(as we encode OTP symmetric key encryption) that I'm not immediately
seeing. Concurrency would be a much bigger lift- though this could
perhaps be addressed using techniques in \cite{MITCHELL2006118}.  I
note that \cite{barthe2019probabilistic} also provides no support for
concurrent reasoning, and they were the source for the OT protocol we
used, and honestly I sort of hoping that we would get the same
indulgences from reviewers on these points, especially since we were
avoiding any claims about UC. But I guess Barthe et al.~were targeting
probabilistic reasoning generally, rather than MPC specifically where
these issues are more central.

\section{Bridging Formal and Computational Views}

Ian cites \cite{10.1007/3-540-44929-9_1,MITCHELL2006118} as examples
of ``proving cryptographic security using PL techniques for
traditional cryptography''. I believe the main point here is about
the relation of our work to previous work on bridging the gap between
formal and computational views of security. Since we have similar
aims the relation is clear at a high level.

Technically speaking there are probably some strong contrasts with
these citations explicitly. For example,
in \cite{10.1007/3-540-44929-9_1} the formal view is characterized by
an equivalence relations on expressions that axiomitizes $\{
0 \}_K \simeq \{ 1 \}_{K'}$ based on primitize assumptions of PK
encryption. However, our goal is a lower level analysis that could
conceivably establish these primitive properties of underlying
protocols.

The methods developed in \cite{10.1007/3-540-44929-9_1} are very
sophisticated and perhaps more powerful than needed for the context
we've been working in. For example, they develop a contextual
congruence relation that looks as if it could support reasoning
about UC properties of protocols. But we have definitely been
considering UC to be out of scope (at least initially). On the
other hand, their machinery for concurrent reasoning could be
extremely helpful as disussed above.

\section{Dimensions of Declassification}

I agree it would be interesting to do a ``who what were when''
analysis for MPC. Another very interesting feature of this work
is the partial equivalence relation (PER) model.  Our notion of
``ideal knowledge'' is adapted from the notion of knowledge as defined
in \cite{10.1007/978-3-540-37621-7_9}, and the PER model is clearly
related to this. More generally, any ideal functionality $F$ induces a
PER (an extensional PER, more precisely) I'll call $\mathit{IdealDeclass}$
that characterizes the downgrading/declassification/delimitation policy,
where $=_C$ is low equivalence (up to equivalence of inputs from corrupt
parties C):
$$
\forall \bar{x}_1,\bar{x}_2\ .\ F(\bar{x}_1) =_C F(\bar{x}_1) \iff \bar{x}_1
\ \mathit{IdealDeclass}\ \bar{x}_2  
$$

However, I'll note that our use of joint probability distributions
provides more subtle machinery that could be more useful, perhaps
when integrating subroutines (OT) that enjoy computational hardness
(vs.~purely formal) correctness properties. For example, if an
MPC output $y$ is known, then any $\bar{x}$ with $F(\bar{x}) = y$
is equally likely according to the ``Sabelfeld style'' formulation
of knowledge, whereas we can assign different $\bar{x}$ different
probabilities in a joint probability distribution.

\section{Comparison with Downgrading Policies}

Ian mentions Viaduct \cite{10.1145/3453483.3454074} which arguably
represents the current state-of-the-art in information flow frameworks
with declassition and endorsement. This was missed related work on our
part, but I would argue that it is really not directly related to our
work, for two critical reasons that we are having trouble explaining.

\emph{First}, one general comment about this body of work is that
downgrading policies as such are defined by the declassifications and
endorsements themselves. For example, gradual release \cite{4223226}
is the policy that says only explicitly declassified information is
released, and robust declassification \cite{930133} is a policy that
guarantees that an active adversary can't learn any more information
through declassifications than a passive adversary. Some work shows
how ``intentional'' release policies are derived syntactically from
program declassifications as reviewed in
\cite{10.5555/1662658.1662659}, but these policy expressions are sort
of awkward and course-grained.

In contrast, MPC establishes an ``upper bound'' on knowledge
(specifically the PER $\mathit{IdealDeclass}$ defined above if we are
thinking that way), and our goal is to ensure that declassification
does not overstep that bound. In fact, we have no explicit
``declassify'' construct (though perhaps view assignments aka message
passing are arguably declassifications). Rather, we argue that the
probability mass function (pmf) is an easily understood ``knowledge
metric'', and we aim to compute the actual mass function associated
with the protocol, for a direct comparison of the (ideal) upper bound
and actual (real) bound.

Going back to Viaduct, my understanding is that this system in the robust
declassification school of thought, where labeling and explicit
downgrading and endorsement establish the policy. They also include
``nonmalleability'' as the dual of robust declassification
\cite{10.1145/3133956.3134054}. A main result is that their compiler
can exploit MPC and other protocols in the back end to support the
policy defined by the security annotations in programs. But they never
aim to compare this ``syntactic'' policy against an extensional policy
(as we do). Another point is that they don't need to consider multiple
partitionings of corrupt and honest parties as is needed in MPC.

\emph{Second}, another general difference is that Viaduct for example
assumes correctness of the MPC protocols they compile to. In this way
Viaduct is similar to Wysteria, for example
\cite{rastogi2014wysteria}. Our goal is to establish a general
frameword for (semi)-automated verification of the MPC protocols
themselves.  This is why our mass function formulation, and methods
for iterating through corrupt/honest partitions, for example, are
important technical details of our system.

Unfortunately, I think we were not able to persuade reviewers on this
point because the automated component of the analysis is exponential
and the manual component looks a lot like probabilistic separation
logic, and is manual. Not sure how to remedy that- though formulating
this within some mechanized metatheory, as a reviewer suggested, would
be an obvious way.

\section{Integrity}

This is a bit speculative but since we are working in a space where
ideal functionalities $F$ are deterministic, we can formulate
integrity via the PER $\mathit{IdealEndorse}$ that characterizes the
following endorsement policy, where $=_H$ is high equivalence (up to
equivalence of inputs from corrupt parties H):
$$
\forall \bar{x}_1,\bar{x}_2\ .\ F(\bar{x}_1) =_H F(\bar{x}_1) \iff \bar{x}_1
\ \mathit{IdealEndorse}\ \bar{x}_2  
$$

However, to capture either nonmalleability or robust declassification
we would need to extend our threat model to capture ``attacks'' as
defined in \cite{930133}. Our model is essentially passive. This is
non-trivial but I think can clearly be accomplished.

The potential application to verifiable computing is an interesting
suggestion.  I frankly need to better understand the relevant security
properties, but in any case, I do think it's possible to formulate notions
of both robust declassification and nonmalleability in our
(probabilistic) language setting. And I suspect there are relations to
security for verifiable computing and nonmalleability, as well as
malicious security and robust declassification. In particular I
speculate that NIMO + robust declassification implies malicious
security.

\section{Generalizations to Realistic Settings}

I need to spend some time understanding the implementation of SPDZ
with SHE that Ian cites. However as a general comment I think our
approach is modular- for example if we replaced our OT primitive where
we can express its information-leakage characteristics as a pmf, we
can drop that directly into our framework. Anyway we need to reflect
on this more, especially related to the main critiques of
oversimplification especially wrt OT.

\bibliographystyle{ACM-Reference-Format}
%\bibliographystyle{alpha}
\bibliography{logic-bibliography,secure-computation-bibliography}

\end{document}
